---
title: "Finding thresholds - Latent Class Analysis (LCA)"
author: ''
subtitle: "MM4DBER" 
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here)
```



# --------------------------------------------------------------------------------------

## Preparation

# --------------------------------------------------------------------------------------

## Data source:

Public-use data, *The Longitudinal Survey of American Youth* (**LSAY**):  [$\color{blue}{\text{See documentation here}}$](https://www.lsay.org/)

# --------------------------------------------------------------------------------------

Load packages
```{r, eval=TRUE}

library(tidyverse)
library(haven)
library(glue)
library(MplusAutomation)
library(here)
library(janitor)
library(gt)
library(reshape2)
library(cowplot) 
library(poLCA)
library(linguisticsdown)

```



```{r, echo=FALSE, eval=TRUE, out.width = "65%", out.height= "65%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/LCA_Diagram_SEMLAB8.png")
```

Load data
```{r}

lsay_data <- read_csv("https://garberadamc.github.io/project-site/data/lca_lsay_sci.csv",
                      na = c("9999", "9999.00")) %>%               
  clean_names() %>%                                                                               
  dplyr::select(1:5, Female=female, Enjoy = ab39m, Useful = ab39t,                                                                  Logical = ab39u, Job = ab39w, Adult = ab39x)                                 


```

View LCA indicators 
```{r, echo=FALSE, eval=TRUE}

var_table <- tribble(
   ~"Name",      ~"Label",  ~"Values",                                   
#--------------|--------------------------------|-----|,
  "Enjoy",   "I enjoy science "                   ,  "0 = Disagree, 1 = Agree",
  "Useful",  "Science useful in everday problems ",  "0 = Disagree, 1 = Agree",
  "Logical", "Science helps logical thinkng "     ,  "0 = Disagree, 1 = Agree",
  "Job",     "Need science for a good job "       ,  "0 = Disagree, 1 = Agree",
  "Adult",   "Will use science often as an adult ",  "0 = Disagree, 1 = Agree",
  "Female",   "Reported gender"                   ,  "0 = Male, 1 = Female")

gt(var_table) %>% 
  tab_header(
    title = "Science Attitudes",  # Add a title
    subtitle = md("&nbsp;")) %>% 
  tab_options(
    table.width = pct(80),
    column_labels.font.weight = "bold"
  ) %>%
  tab_footnote(
    footnote = "Longitudinal Study of American Youth",
    location = cells_title()) %>%
  tab_row_group(
    group = "LCA Indicators",
    rows = 1:5) %>%
  tab_row_group(
    group = "Covariate",
    rows = 6:6) %>% 
  row_group_order(
      groups = c("LCA Indicators","Covariate")
      )


```

# --------------------------------------------------------------------------------------

##  2.1 Enumeration: 

Estimate *K*-class models with 1 through 6 classes.

# --------------------------------------------------------------------------------------

Run enumeration using `mplusObject` method
```{r, eval=FALSE}

lca_k1_6  <- lapply(1:6, function(k) {
  
  lca_enum  <- mplusObject(
      
    TITLE = glue("Class {k}"), 
  
    VARIABLE = glue(
    "categorical = Enjoy-Adult; 
     usevar = Enjoy-Adult;
     classes = c({k}); "),
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture;
    starts = 200 100; 
    processors = 10;",
  
  OUTPUT = "sampstat residual tech11 tech14;",
  
  PLOT = 
    "type = plot3; 
    series = Enjoy-Adult(*);",
  
  usevariables = colnames(lsay_data),
  rdata = lsay_data)

lca_enum_fit <- mplusModeler(lca_enum, 
                            dataout=glue(here("enum_lsay", "lca_lsay.dat")),
                            modelout=glue(here("enum_lsay", "c{k}_lca.inp")) ,
                            check=TRUE, run = TRUE, hashfilename = FALSE)
})

```

# --------------------------------------------------------------------------------------

## Plot Final Model - Conditional Item Probability Plot 

# --------------------------------------------------------------------------------------

### This syntax creates a function called `plot_lca_function` that requires 7 arguments (inputs):

- `model_name`: name of Mplus model object (e.g., `model_step1`)
- `item_num`: the number of items in LCA measurement model (e.g., `5`)
- `class_num`: the number of classes (*k*) in LCA model (e.g., `3`)
- `item_labels`: the item labels for x-axis (e.g., `c("Enjoy","Useful","Logical","Job","Adult")`)
- `class_labels`: the class label names (e.g., `c("Adaptive Coping","Externalizing Behavior","No Coping")`)
- `class_legend_order` = change the order that class names are listed in the plot legend (e.g., `c(2,1,3)`)
- `plot_title`: include the title of the plot here (e.g., `"LCA Posterior Probability Plot"`)

```{r}

plot_lca_function <- function(model_name,item_num,class_num,item_labels,
                              class_labels,class_legend_order,plot_title){

mplus_model <- as.data.frame(model_name$gh5$means_and_variances_data$estimated_probs$values)
plot_data <- mplus_model[seq(2, 2*item_num, 2),]

c_size <- as.data.frame(model_name$class_counts$modelEstimated$proportion)
colnames(c_size) <- paste0("cs")
c_size <- c_size %>% mutate(cs = round(cs*100, 2))
colnames(plot_data) <- paste0(class_labels, glue(" ({c_size[1:class_num,]}%)"))
plot_data <- plot_data %>% relocate(class_legend_order)

plot_data <- cbind(Var = paste0("U", 1:item_num), plot_data)
plot_data$Var <- factor(plot_data$Var,
               labels = item_labels)
plot_data$Var <- fct_inorder(plot_data$Var)

pd_long_data <- melt(plot_data, id.vars = "Var") 

# This syntax uses the data.frame created above to produce the plot with `ggplot()`

p <- pd_long_data %>%
  ggplot(aes(x = as.integer(Var), y = value,
  shape = variable, colour = variable, lty = variable)) +
  geom_point(size = 4) + geom_line() + 
  scale_x_continuous("", breaks = 1:item_num, labels = plot_data$Var) + 
  labs(title = plot_title, y = "Probability") +
  theme_cowplot() +
  theme(legend.title = element_blank(), 
        legend.position = "top")

p
return(p)
}

```

```{r}

model_c4 <- readModels(here("enum_lsay", "c4_lca.out"), quiet = TRUE)
                           
```

## Run the `plot_lca_function` by specifying each input (*Figure 1*)
```{r}

plot_lca_function(
  model_name = model_c4, 
  item_num = 5,
  class_num = 4,
  item_labels = c("Enjoy", "Useful", "Logical", "Job", "Adult"),
  class_labels = c("Class1","Class2","Class3","Class4"),
  class_legend_order = c(1,2,3,4),
  plot_title = "Science Attitudes"
  )

```
```{r}
ggsave(here("figures","C4_LSAY_Plot.png"), dpi=300, height=4, width=6, units="in")
```

Run model with covariate
```{r, eval=FALSE}


  lca_enum  <- mplusObject(
      
    TITLE = "Class 4 model with covariate (female)", 
  
    VARIABLE = 
    "categorical = Enjoy-Adult; 
     usevar = Enjoy-Adult;
     
     classes = c(4); 
     auxiliary = Female (r3step);",
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture;
    starts = 500 200; 
    processors = 10;",
  
  OUTPUT = "sampstat residual tech11 tech14;",
  
  PLOT = 
    "type = plot3; 
    series = Enjoy-Adult(*);",
  
  usevariables = colnames(lsay_data),
  rdata = lsay_data)

lca_enum_fit <- mplusModeler(lca_enum, 
                            dataout=here("cov_c4", "lca_lsay.dat"),
                            modelout=here("cov_c4", "c4_lca_cov.inp") ,
                            check=TRUE, run = TRUE, hashfilename = FALSE)

```




